---
title: "Punctate Report"
author: "David McGaughey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook:
    theme: flatly
    toc: true
---

# Identifying genes with the following characteristics:
- 2 or more sum ATAC-peaks in GFP versus iPSC in the 1000bp and 5000bp upstream windows
- Changes in RNA expression between GFP and iPSC of abs(log2(TPM+1)) > 2

```{r, fig.height=3.3, fig.width=4, warning = F , message = F}
library(tidyverse)
library(ComplexHeatmap)
library(viridis)
library(circlize)
library(annotables)
#peak <- read_tsv('/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/peak_full/homer_GFP__not__IPSC/all/all_common_peaks.blackListed.narrowPeak.closestTSS__interesting_homer_motif.bed.gz')
expression <- read_csv('~/git/ipsc_rpe_RNA-seq/data/lsTPM_by_Line_with_Diff_Exp.tsv', col_types = 'ccdddcddd')

# iPSC
cgm_ipsc_value <- read_tsv('/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/CGM/IPSC_ATAC-Seq_common_peaks.blackListed.narrowPeak.CGM_score.tsv', col_types = 'ccddddddddddddddddd') %>% 
  left_join(. , expression %>% filter(Line == 'iPSC') %>% select(Gene, lsTPM), by = c('Gene_Name' = 'Gene'))
cgm_ipsc_score <- read_tsv('/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/CGM/IPSC_ATAC-Seq_common_peaks.blackListed.narrowPeak.CGM_score.tsv', col_types = 'ccddddddddddddddddd') %>% 
  left_join(. , expression %>% filter(Line == 'iPSC') %>% select(Gene, lsTPM), by = c('Gene_Name' = 'Gene'))

# GFP
cgm_gfp_value <- read_tsv('/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/CGM/GFP_ATAC-Seq_common_peaks.blackListed.narrowPeak.CGM_score.tsv', col_types = 'ccddddddddddddddddd') %>% 
  left_join(. , expression %>% filter(Line == 'GFP') %>% select(Gene, lsTPM), by = c('Gene_Name' = 'Gene'))
cgm_gfp_score <- read_tsv('/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/CGM/GFP_ATAC-Seq_common_peaks.blackListed.narrowPeak.CGM_score.tsv', col_types = 'ccddddddddddddddddd') %>% 
  left_join(. , expression %>% filter(Line == 'GFP') %>% select(Gene, lsTPM), by = c('Gene_Name' = 'Gene'))
cgm_gfp_h3k27ac_value <- read_tsv('/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/CGM/GFP_ATAC-Seq_common_peaks.h3k27ac_intersect.blackListed.narrowPeak.CGM_score.tsv', col_types = 'ccddddddddddddddddd') %>% 
  left_join(. , expression %>% filter(Line == 'GFP') %>% select(Gene, lsTPM), by = c('Gene_Name' = 'Gene'))
cgm_gfp_h3k27ac_score <- read_tsv('/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/CGM/GFP_ATAC-Seq_common_peaks.h3k27ac_intersect.blackListed.narrowPeak.CGM_score.tsv', col_types = 'ccddddddddddddddddd') %>% 
  left_join(. , expression %>% filter(Line == 'GFP') %>% select(Gene, lsTPM), by = c('Gene_Name' = 'Gene'))

# RFP
cgm_rfp_value <- read_tsv('/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/CGM/RFP_ATAC-Seq_common_peaks.blackListed.narrowPeak.CGM_score.tsv', col_types = 'ccddddddddddddddddd') %>% 
  left_join(. , expression %>% filter(Line == 'RFP') %>% select(Gene, lsTPM), by = c('Gene_Name' = 'Gene'))
cgm_rfp_score <- read_tsv('/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/CGM/RFP_ATAC-Seq_common_peaks.blackListed.narrowPeak.CGM_score.tsv', col_types = 'ccddddddddddddddddd') %>% 
  left_join(. , expression %>% filter(Line == 'RFP') %>% select(Gene, lsTPM), by = c('Gene_Name' = 'Gene'))
cgm_rfp_h3k27ac_value <- read_tsv('/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/CGM/RFP_ATAC-Seq_common_peaks.h3k27ac_intersect.blackListed.narrowPeak.CGM_score.tsv', col_types = 'ccddddddddddddddddd') %>% 
  left_join(. , expression %>% filter(Line == 'RFP') %>% select(Gene, lsTPM), by = c('Gene_Name' = 'Gene'))
cgm_rfp_h3k27ac_score <- read_tsv('/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/CGM/RFP_ATAC-Seq_common_peaks.h3k27ac_intersect.blackListed.narrowPeak.CGM_score.tsv', col_types = 'ccddddddddddddddddd') %>% 
  left_join(. , expression %>% filter(Line == 'RFP') %>% select(Gene, lsTPM), by = c('Gene_Name' = 'Gene'))
```

# iPSC -> GFP
ATAC + RNA-seq

Not using the h3k27ac intersection for this comparison, as the iPSC h3k27ac ChIP-seq samples I found had horrific (near zero) overlap with our iPSC ATAC data.

Narrowing in on promoter region (1000bp) with ATAC score > 100 (GSX2 has a peak score of 105 within 1000bp of TSS)

## GFP - iPSC: Higher Score in TSS (> 100), Higher gene expression (>=2X or more GFP/iPSC)
```{r, fig.height=11, fig.width=4}
delta_cgm <- (cgm_gfp_score %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>%  select(Exon_Value:Gene_length, `log2(TPM+1)`)) - 
  (cgm_ipsc_score %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>% select(Exon_Value:Gene_length, `log2(TPM+1)`)) %>% as.tibble()
delta_cgm$Gene_Name = cgm_gfp_score$Gene_Name
#delta_cgm$var <- apply(delta_cgm %>% select(Exon_Value:`-1e6`), 1, var)

punctate_upstream <- delta_cgm %>% #mutate(`5k_upstream` = `1000`) %>% 
  filter(`1000` > 100, `log2(TPM+1)` > 1) %>% unique() %>% 
  arrange(`log2(TPM+1)`) 
print(nrow(punctate_upstream))
punctate_upstream %>% left_join(., grch38 %>% select(symbol, description), by = c("Gene_Name" = "symbol")) %>% 
  select(Gene_Name, description, `log2(TPM+1)`, Exon_Value:`-1e6`) %>% unique() %>% 
  arrange(`log2(TPM+1)`) %>% 
  DT::datatable()

# output
write(punctate_upstream %>% 
            left_join(., grch38 %>% select(ensgene, symbol, description), by = c("Gene_Name" = "symbol")) %>% 
            left_join(grch38_tx2gene) %>% 
            pull(enstxp) %>% 
            unique(),
          file = '/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/punctate/punctate_analysis_1.txt')


row.names(punctate_upstream) <- punctate_upstream$Gene_Name  
punctate_upstream <- punctate_upstream %>% select(`1e6`:`-1e6`, `log2(TPM+1)`)
atac_heatmap <- punctate_upstream %>% select(-`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-200, 0, 200), viridis(3)),
          cluster_columns = F,
          cluster_rows = F,
          name = 'Peak Score in Window GFP - iPSC',
          show_row_names = F)
expression_heatmap <- punctate_upstream %>% select(`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-3, 0, 3), magma(3)),
          cluster_columns = F,
          name = 'log2(TPM+1) Delta GFP - iPSC')
atac_heatmap + expression_heatmap
```

## GFP - iPSC: Higher Score in TSS (> 100), less gene expression (-2X or more GFP/iPSC)
```{r, fig.height=6, fig.width=4}
delta_cgm <- (cgm_gfp_score %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>%  select(Exon_Value:Gene_length, `log2(TPM+1)`)) - 
  (cgm_ipsc_score %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>% select(Exon_Value:Gene_length, `log2(TPM+1)`)) %>% as.tibble()
delta_cgm$Gene_Name = cgm_gfp_score$Gene_Name
#delta_cgm$var <- apply(delta_cgm %>% select(Exon_Value:`-1e6`), 1, var)

punctate_upstream <- delta_cgm %>% #mutate(`5k_upstream` = `1000`) %>% 
  filter(`1000` > 100, `log2(TPM+1)` < -1) %>% unique() %>% 
  arrange(`log2(TPM+1)`) 
print(nrow(punctate_upstream))
punctate_upstream %>% left_join(., grch38 %>% select(symbol, description), by = c("Gene_Name" = "symbol")) %>% 
  select(Gene_Name, description, `log2(TPM+1)`, Exon_Value:`-1e6`) %>% unique() %>% 
  arrange(`log2(TPM+1)`) %>% 
  DT::datatable()

# output
write(punctate_upstream %>% 
            left_join(., grch38 %>% select(ensgene, symbol, description), by = c("Gene_Name" = "symbol")) %>% 
            left_join(grch38_tx2gene) %>% 
            pull(enstxp) %>% 
            unique(),
          file = '/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/punctate/punctate_analysis_2.txt')


row.names(punctate_upstream) <- punctate_upstream$Gene_Name  
punctate_upstream <- punctate_upstream %>% select(`1e6`:`-1e6`, `log2(TPM+1)`)
atac_heatmap <- punctate_upstream %>% select(-`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-200, 0, 200), viridis(3)),
          cluster_columns = F,
          cluster_rows = F,
          name = 'Peak Score in Window GFP - iPSC',
          show_row_names = F)
expression_heatmap <- punctate_upstream %>% select(`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-3, 0, 3), magma(3)),
          cluster_columns = F,
          name = 'log2(TPM+1) Delta GFP - iPSC')
atac_heatmap + expression_heatmap
```

## GFP - iPSC: Lower Score in TSS (< -100), higher gene expression (2X or more GFP/iPSC)
```{r, fig.height=9, fig.width=4}
delta_cgm <- (cgm_gfp_score %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>%  select(Exon_Value:Gene_length, `log2(TPM+1)`)) - 
  (cgm_ipsc_score %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>% select(Exon_Value:Gene_length, `log2(TPM+1)`)) %>% as.tibble()
delta_cgm$Gene_Name = cgm_gfp_score$Gene_Name
#delta_cgm$var <- apply(delta_cgm %>% select(Exon_Value:`-1e6`), 1, var)

punctate_upstream <- delta_cgm %>% #mutate(`5k_upstream` = `1000`) %>% 
  filter(`1000` < -100, `log2(TPM+1)` > 1) %>% unique() %>% 
  arrange(`log2(TPM+1)`) 
print(nrow(punctate_upstream))
punctate_upstream %>% left_join(., grch38 %>% select(symbol, description), by = c("Gene_Name" = "symbol")) %>% 
  select(Gene_Name, description, `log2(TPM+1)`, Exon_Value:`-1e6`) %>% unique() %>% 
  arrange(`log2(TPM+1)`) %>% 
  DT::datatable()

# output
write(punctate_upstream %>% 
            left_join(., grch38 %>% select(ensgene, symbol, description), by = c("Gene_Name" = "symbol")) %>% 
            left_join(grch38_tx2gene) %>% 
            pull(enstxp) %>% 
            unique(), 
          file = '/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/punctate/punctate_analysis_3.txt')


row.names(punctate_upstream) <- punctate_upstream$Gene_Name  
punctate_upstream <- punctate_upstream %>% select(`1e6`:`-1e6`, `log2(TPM+1)`)
atac_heatmap <- punctate_upstream %>% select(-`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-200, 0, 200), viridis(3)),
          cluster_columns = F,
          cluster_rows = F,
          name = 'Peak Score in Window GFP - iPSC',
          show_row_names = F)
expression_heatmap <- punctate_upstream %>% select(`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-3, 0, 3), magma(3)),
          cluster_columns = F,
          name = 'log2(TPM+1) Delta GFP - iPSC')
atac_heatmap + expression_heatmap
```

## GFP - iPSC: Lower Score in TSS (< -100), lower gene expression (-2X or more GFP/iPSC)
```{r, fig.height=12, fig.width=4}
delta_cgm <- (cgm_gfp_score %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>%  select(Exon_Value:Gene_length, `log2(TPM+1)`)) - 
  (cgm_ipsc_score %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>% select(Exon_Value:Gene_length, `log2(TPM+1)`)) %>% as.tibble()
delta_cgm$Gene_Name = cgm_gfp_score$Gene_Name
#delta_cgm$var <- apply(delta_cgm %>% select(Exon_Value:`-1e6`), 1, var)

punctate_upstream <- delta_cgm %>% #mutate(`5k_upstream` = `1000`) %>% 
  filter(`1000` < -100, `log2(TPM+1)` < -1) %>% unique() %>% 
  arrange(`log2(TPM+1)`) 
print(nrow(punctate_upstream))
punctate_upstream %>% left_join(., grch38 %>% select(symbol, description), by = c("Gene_Name" = "symbol")) %>% 
  select(Gene_Name, description, `log2(TPM+1)`, Exon_Value:`-1e6`) %>% unique() %>% 
  arrange(`log2(TPM+1)`) %>% 
  DT::datatable()

# output
write(punctate_upstream %>% 
            left_join(., grch38 %>% select(ensgene, symbol, description), by = c("Gene_Name" = "symbol")) %>% 
            left_join(grch38_tx2gene) %>% 
            pull(enstxp) %>% 
            unique(), 
          file = '/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/punctate/punctate_analysis_4.txt')


row.names(punctate_upstream) <- punctate_upstream$Gene_Name  
punctate_upstream <- punctate_upstream %>% select(`1e6`:`-1e6`, `log2(TPM+1)`)
atac_heatmap <- punctate_upstream %>% select(-`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-200, 0, 200), viridis(3)),
          cluster_columns = F,
          cluster_rows = F,
          name = 'Peak Score in Window GFP - iPSC',
          show_row_names = F)
expression_heatmap <- punctate_upstream %>% select(`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-3, 0, 3), magma(3)),
          cluster_columns = F,
          name = 'log2(TPM+1) Delta GFP - iPSC')
atac_heatmap + expression_heatmap
```


# GFP vs RFP

Modified approach from `trends.Rmd` where the filtering is based upon evaluating two settings:
  - Within 1000bp
  - Within 100,000bp
  - (Not using the entire region)

## GFP - RFP: Higher Peak Score (1000bp > 30 or 100k > 200) and Higher Expression (>0.5 log2(TPM) GFP - RFP)
**WITH h3k27ac Intersection**

0.5 log(TPM+1) delta set as GSX2 is just over 0.5 log2(TPM+1) overexpressed GFP - RFP
```{r, fig.height=6, fig.width=4}
delta_cgm <- (cgm_gfp_h3k27ac_value %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>%  select(Exon_Value:Gene_length, `log2(TPM+1)`)) - 
             (cgm_rfp_h3k27ac_value %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>% select(Exon_Value:Gene_length, `log2(TPM+1)`)) %>% as.tibble()
delta_cgm$Gene_Name = cgm_gfp_h3k27ac_score$Gene_Name
#delta_cgm$var <- apply(delta_cgm %>% select(Exon_Value:`-1e6`), 1, var)

delta_cgm <- delta_cgm %>% mutate(Promoter = `1000`,
                                  Upstream100K = `1000` + `5000` + `1e4` + `5e4` + `1e5`) %>% 
  unique()
#delta_cgm %>% select(Exon_Value:`-1e6`) %>% rowSums(na.rm=TRUE) -> delta_cgm$All


punctate_upstream <- delta_cgm %>% 
  filter(`log2(TPM+1)` > 0.5) %>% filter(Promoter > 30 | Upstream100K > 200) %>% 
  arrange(`log2(TPM+1)`)
punctate_upstream %>% left_join(., grch38 %>% select(symbol, description), by = c("Gene_Name" = "symbol")) %>% 
  select(Gene_Name, description, `log2(TPM+1)`, Promoter, Upstream100K, Exon_Value:`-1e6`) %>% 
  DT::datatable()

# output
write(punctate_upstream %>% 
            left_join(., grch38 %>% select(ensgene, symbol, description), by = c("Gene_Name" = "symbol")) %>% 
            left_join(grch38_tx2gene) %>% 
            pull(enstxp) %>% 
            unique(), 
          file = '/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/punctate/punctate_analysis_5.txt')


row.names(punctate_upstream) <- punctate_upstream$Gene_Name  
punctate_upstream <- punctate_upstream %>% select(`1e6`:`-1e6`, `log2(TPM+1)`, Promoter, Upstream100K)
atac_heatmap <- punctate_upstream %>% select(-`log2(TPM+1)`, -Promoter, -Upstream100K) %>% 
  Heatmap(col = colorRamp2(c(-100, 0, 100), viridis(3)),
          cluster_columns = F,
          name = 'Peaks Score in Window GFP - RFP',
          cluster_rows = F,
          show_row_names = F)

region_heatmap <- punctate_upstream %>% select(Promoter, Upstream100K) %>% 
  Heatmap(col = colorRamp2(c(-100, 0, 100), viridis(3)),
          cluster_columns = F,
          name = 'Aggregation Peak Score GFP - RFP',
          show_row_names = F,
          cluster_rows = F,
          show_heatmap_legend = F)

expression_heatmap <- punctate_upstream %>% select(`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-2, 0, 2), magma(3)),
          cluster_columns = F,
          cluster_rows = F,
          name = 'log2(TPM+1) Delta GFP - RFP')
atac_heatmap + region_heatmap + expression_heatmap

```

## GFP - RFP: Higher Peak Score (1000bp > 30 or 100k > 200) and Lower Expression (< -0.5 log2(TPM) GFP - RFP)
Notice how there are fewer...which is good, as ATAC + h3k27Ac is a canonical "activation" signal
```{r, fig.height=3, fig.width=4}
delta_cgm <- (cgm_gfp_h3k27ac_value %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>%  select(Exon_Value:Gene_length, `log2(TPM+1)`)) - 
             (cgm_rfp_h3k27ac_value %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>% select(Exon_Value:Gene_length, `log2(TPM+1)`)) %>% as.tibble()
delta_cgm$Gene_Name = cgm_gfp_h3k27ac_value$Gene_Name
#delta_cgm$var <- apply(delta_cgm %>% select(Exon_Value:`-1e6`), 1, var)

delta_cgm <- delta_cgm %>% mutate(Promoter = `1000`,
                                  Upstream100K = `1000` + `5000` + `1e4` + `5e4` + `1e5`) %>% 
  unique()
#delta_cgm %>% select(Exon_Value:`-1e6`) %>% rowSums(na.rm=TRUE) -> delta_cgm$All


punctate_upstream <- delta_cgm %>% 
  filter(`log2(TPM+1)` < -0.5) %>% filter(Promoter > 30 | Upstream100K > 200) %>% 
  arrange(`log2(TPM+1)`)
punctate_upstream %>% left_join(., grch38 %>% select(symbol, description), by = c("Gene_Name" = "symbol")) %>% 
  select(Gene_Name, description, `log2(TPM+1)`, Promoter, Upstream100K, Exon_Value:`-1e6`) %>% 
  DT::datatable()

# output
write(punctate_upstream %>% 
            left_join(., grch38 %>% select(ensgene, symbol, description), by = c("Gene_Name" = "symbol")) %>% 
            left_join(grch38_tx2gene) %>% 
            pull(enstxp) %>% 
            unique(), 
          file = '/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/punctate/punctate_analysis_6.txt')


row.names(punctate_upstream) <- punctate_upstream$Gene_Name  
punctate_upstream <- punctate_upstream %>% select(`1e6`:`-1e6`, `log2(TPM+1)`, Promoter, Upstream100K)
atac_heatmap <- punctate_upstream %>% select(-`log2(TPM+1)`, -Promoter, -Upstream100K) %>% 
  Heatmap(col = colorRamp2(c(-100, 0, 100), viridis(3)),
          cluster_columns = F,
          name = 'Peaks Score in Window GFP - RFP',
          cluster_rows = F,
          show_row_names = F)

region_heatmap <- punctate_upstream %>% select(Promoter, Upstream100K) %>% 
  Heatmap(col = colorRamp2(c(-100, 0, 100), viridis(3)),
          cluster_columns = F,
          name = 'Aggregation Peak Score GFP - RFP',
          show_row_names = F,
          cluster_rows = F,
          show_heatmap_legend = F)

expression_heatmap <- punctate_upstream %>% select(`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-2, 0, 2), magma(3)),
          cluster_columns = F,
          cluster_rows = F,
          name = 'log2(TPM+1) Delta GFP - RFP')
atac_heatmap + region_heatmap + expression_heatmap

```

## GFP - RFP: Lower Peak Score (1000bp < -30 or 100k < -200) and Higher Expression (>1 log2(TPM) GFP - RFP)

I've increased the stringency for log2(TPM+1) change from 0.5 to 1, as there are way too many results.
```{r, fig.height=6, fig.width=4}
delta_cgm <- (cgm_gfp_h3k27ac_value %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>%  select(Exon_Value:Gene_length, `log2(TPM+1)`)) - 
             (cgm_rfp_h3k27ac_value %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>% select(Exon_Value:Gene_length, `log2(TPM+1)`)) %>% as.tibble()
delta_cgm$Gene_Name = cgm_gfp_h3k27ac_value$Gene_Name
#delta_cgm$var <- apply(delta_cgm %>% select(Exon_Value:`-1e6`), 1, var)

delta_cgm <- delta_cgm %>% mutate(Promoter = `1000`,
                                  Upstream100K = `1000` + `5000` + `1e4` + `5e4` + `1e5`) %>% 
  unique()
#delta_cgm %>% select(Exon_Value:`-1e6`) %>% rowSums(na.rm=TRUE) -> delta_cgm$All


punctate_upstream <- delta_cgm %>% 
  filter(`log2(TPM+1)` > 1) %>% filter(Promoter < -30 | Upstream100K < -200) %>% 
  arrange(`log2(TPM+1)`)
punctate_upstream %>% left_join(., grch38 %>% select(symbol, description), by = c("Gene_Name" = "symbol")) %>% 
  select(Gene_Name, description, `log2(TPM+1)`, Promoter, Upstream100K, Exon_Value:`-1e6`) %>% 
  DT::datatable()

# output
write(punctate_upstream %>% 
            left_join(., grch38 %>% select(ensgene, symbol, description), by = c("Gene_Name" = "symbol")) %>% 
            left_join(grch38_tx2gene) %>% 
            pull(enstxp) %>% 
            unique(), 
          file = '/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/punctate/punctate_analysis_7.txt')


row.names(punctate_upstream) <- punctate_upstream$Gene_Name  
punctate_upstream <- punctate_upstream %>% select(`1e6`:`-1e6`, `log2(TPM+1)`, Promoter, Upstream100K)
atac_heatmap <- punctate_upstream %>% select(-`log2(TPM+1)`, -Promoter, -Upstream100K) %>% 
  Heatmap(col = colorRamp2(c(-100, 0, 100), viridis(3)),
          cluster_columns = F,
          name = 'Peaks Score in Window GFP - RFP',
          cluster_rows = F,
          show_row_names = F)

region_heatmap <- punctate_upstream %>% select(Promoter, Upstream100K) %>% 
  Heatmap(col = colorRamp2(c(-100, 0, 100), viridis(3)),
          cluster_columns = F,
          name = 'Aggregation Peak Score GFP - RFP',
          show_row_names = F,
          cluster_rows = F,
          show_heatmap_legend = F)

expression_heatmap <- punctate_upstream %>% select(`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-2, 0, 2), magma(3)),
          cluster_columns = F,
          cluster_rows = F,
          name = 'log2(TPM+1) Delta GFP - RFP')
atac_heatmap + region_heatmap + expression_heatmap

```

## GFP - RFP: Lower Peak Score (1000bp < -30 or 100k < -200) and Lower Expression (< -1 log2(TPM) GFP - RFP)

I've increased the stringency for log2(TPM+1) change from 0.5 to 1, as there are way too many results.
```{r, fig.height=3, fig.width=4}
delta_cgm <- (cgm_gfp_h3k27ac_value %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>%  select(Exon_Value:Gene_length, `log2(TPM+1)`)) - 
             (cgm_rfp_h3k27ac_value %>% mutate(`log2(TPM+1)` = log2(lsTPM+1)) %>% select(Exon_Value:Gene_length, `log2(TPM+1)`)) %>% as.tibble()
delta_cgm$Gene_Name = cgm_gfp_h3k27ac_value$Gene_Name
#delta_cgm$var <- apply(delta_cgm %>% select(Exon_Value:`-1e6`), 1, var)

delta_cgm <- delta_cgm %>% mutate(Promoter = `1000`,
                                  Upstream100K = `1000` + `5000` + `1e4` + `5e4` + `1e5`) %>% 
  unique()
#delta_cgm %>% select(Exon_Value:`-1e6`) %>% rowSums(na.rm=TRUE) -> delta_cgm$All


punctate_upstream <- delta_cgm %>% 
  filter(`log2(TPM+1)` < -1) %>% filter(Promoter < -30 | Upstream100K < -200) %>% 
  arrange(`log2(TPM+1)`)
punctate_upstream %>% left_join(., grch38 %>% select(symbol, description), by = c("Gene_Name" = "symbol")) %>% 
  select(Gene_Name, description, `log2(TPM+1)`, Promoter, Upstream100K, Exon_Value:`-1e6`) %>% 
  DT::datatable()

# output
write(punctate_upstream %>% 
            left_join(., grch38 %>% select(ensgene, symbol, description), by = c("Gene_Name" = "symbol")) %>% 
            left_join(grch38_tx2gene) %>% 
            pull(enstxp) %>% 
            unique() %>% 
          file = '/data/mcgaugheyd/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/punctate/punctate_analysis_8.txt')


row.names(punctate_upstream) <- punctate_upstream$Gene_Name  
punctate_upstream <- punctate_upstream %>% select(`1e6`:`-1e6`, `log2(TPM+1)`, Promoter, Upstream100K)
atac_heatmap <- punctate_upstream %>% select(-`log2(TPM+1)`, -Promoter, -Upstream100K) %>% 
  Heatmap(col = colorRamp2(c(-100, 0, 100), viridis(3)),
          cluster_columns = F,
          name = 'Peaks Score in Window GFP - RFP',
          cluster_rows = F,
          show_row_names = F)

region_heatmap <- punctate_upstream %>% select(Promoter, Upstream100K) %>% 
  Heatmap(col = colorRamp2(c(-100, 0, 100), viridis(3)),
          cluster_columns = F,
          name = 'Aggregation Peak Score GFP - RFP',
          show_row_names = F,
          cluster_rows = F,
          show_heatmap_legend = F)

expression_heatmap <- punctate_upstream %>% select(`log2(TPM+1)`) %>% 
  Heatmap(col = colorRamp2(c(-2, 0, 2), magma(3)),
          cluster_columns = F,
          cluster_rows = F,
          name = 'log2(TPM+1) Delta GFP - RFP')
atac_heatmap + region_heatmap + expression_heatmap

```
