---
title: Motif / TFBS Analysis
author: David McGaughey
date: '`r format(Sys.Date(), "%Y-%m-%d")`'
output: 
  html_notebook:
    theme: flatly
    toc: true
    code_folding: hide
---

# Workflow to ID motifs and match to genes

(Full implementation in Snakefile)

1. Download TF motifs from cisbp
2. Only keep TF which are abs(logFc) > 1 between GFP/RPE and iPSC (~1200)
3. Check for motifs in narrow ATAC-seq peaks with fimo
4. Identify closest 10 transcripts to each motif
5. Bootstrap steps 2 and 3 100 times to get background rate

```{r, message=F, warning=F, results='hide'}
# Load Libraries without printing any warnings or messages
library(tidyverse)
library(ggridges)
library(cowplot)
```

# Motif/TF assessed
```{r, results = 'hide'}
# Load in real data
samples_full = list.files('/Volumes/data/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/fimo_motifs/', pattern = '*dat.gz', full.names = T)

fimo_data <- list()
for (i in samples_full){
  sample <- str_split(str_split(i, '/')[[1]][10], '\\.')[[1]][1]
  input <- read_tsv(i)
  input <- input %>% mutate(sequence_name = as.character(sequence_name)) %>% mutate(sample = sample)
  fimo_data[[sample]] <- input
}
sample_motifs = bind_rows(fimo_data) %>% mutate(sample=as.factor(sample))

#sample_motifs %>% group_by(sample, motif_alt_id) %>% summarise(Count=n()) %>% arrange(-Count)

tf_motif <- sample_motifs %>% select('TF' = `# motif_id`, motif_alt_id) %>% unique()
tf_motif %>% DT::datatable()
```


```{r, warning=F, message=F, echo=F, results='hide'}
# Load in bootstraps
bootstrap_stats = list.files('/Volumes/data/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/fimo_motifs/bootstrapping_stats/', pattern = '*dat.gz', full.names = T)
#samples_full = list.files('~/Desktop/fimo_motifs/bootstrapping/', pattern = '*dat.gz', full.names = T)

# big difference here vs above
# not saving full file, as each bootstrap is ~200mb COMPRESSED
# just saving the stats
# right now, counts by motif/sample
fimo_data <- list()
for (i in bootstrap_stats){
  sample <- str_split(str_split(i, '/')[[1]][11], '\\.')[[1]][1]
  bootstrap_num <- str_split(str_split(str_split(i, '/')[[1]][11], '\\.')[[1]][2], '_')[[1]][2]
  sample_boot <- paste0(sample,'_',bootstrap_num)
  stats <- read_tsv(i) %>% mutate(sample = as.factor(sample), bootstrap = bootstrap_num)
  fimo_data[[sample_boot]] <- stats
}
bootstrap_counts = bind_rows(fimo_data) %>% mutate(sample=as.factor(sample)) %>% left_join(tf_motif)
```

```{r}
# merge sample counts with bootstrap counts
all <- bind_rows(sample_motifs %>% group_by(sample, motif_alt_id) %>% summarise(Count=n()) %>% ungroup() %>% mutate(bootstrap = 'real'),
                 bootstrap_counts)

# motif ggridges plotter

plotter <- function(motif, scale=TRUE) {
  if (scale == TRUE){
    all <- all %>% filter(motif_alt_id == !!motif) %>% 
      group_by(sample, motif_alt_id) %>% 
      mutate(`Z score` = scale(Count)) %>% 
      ungroup() 
    all %>% 
      mutate(sample = factor(sample, levels=c('iPSC_IIi_9','iPSC_IIJ_10','iPSC_IIK_11','iPSC_IIL_12','RFP_IIE_1','RFP_IIF_2','RFP_IIG_3','RFP_IIH_4','GFP_IIE_5','GFP_IIF_6','GFP_IIG_7','GFP_IIH_8'))) %>% 
      ggplot(aes(x=`Z score`, y=sample)) + 
      geom_density_ridges() +
      geom_point(data = all %>% 
                   filter(bootstrap == 'real', motif_alt_id == !!motif) %>% 
                   ungroup(), aes(x=`Z score`,y=sample), colour='blue', size=2, alpha=0.5) +
      scale_y_discrete(expand = c(0.01, 0)) +
      theme_ridges() + 
      ggtitle(paste0(motif, ' (', 
                     tf_motif %>% filter(motif_alt_id == !!motif) %>% pull(TF),
                     ')'))
  }
  else {
    all %>% filter(motif_alt_id == !!motif) %>% 
      mutate(sample = factor(sample, levels=c('iPSC_IIi_9','iPSC_IIJ_10','iPSC_IIK_11','iPSC_IIL_12','RFP_IIE_1','RFP_IIF_2','RFP_IIG_3','RFP_IIH_4','GFP_IIE_5','GFP_IIF_6','GFP_IIG_7','GFP_IIH_8'))) %>% 
      ggplot(aes(x=Count, y=sample)) + 
      geom_density_ridges() +
      geom_point(data = sample_motifs %>% 
                   filter(motif_alt_id == !!motif) %>% 
                   group_by(sample, motif_alt_id) %>%
                   summarise(Count=n()) %>% ungroup(), aes(x=Count,y=sample), colour='blue', size=2, alpha=0.5) +
      scale_y_discrete(expand = c(0.01, 0)) +
      theme_ridges() + 
      ggtitle(paste0(motif, ' (', 
                     tf_motif %>% filter(motif_alt_id == !!motif) %>% pull(TF),
                     ')'))
  }
}
```
# Plot enrichment  specific motif/TFBS

## TET1
```{r}
plotter('M0610_1.02')
```

## DNMT1
```{r}
plotter('M0609_1.02')
```

## PAX6
```{r}
plotter('M6410_1.02')
```

## SOX10
```{r}
plotter('M6470_1.02')
```

## MITF
```{r}
plotter('M6345_1.02')
```

## HES1
```{r}
plotter('M6271_1.02')
```

## SIX3
```{r}
plotter('M1016_1.02')
```

## SIX2
```{r}
plotter('M0952_1.02')
```

# Distribution of TFBS/motif enrichment
```{r}
delta_motif_FC <- all %>% mutate(Type = case_when(grepl('GFP', sample) ~ 'GFP',
                                                    grepl('RFP', sample) ~ 'RFP',
                                                    TRUE ~ 'iPSC'),
                                   Boot = case_when(bootstrap == 'real' ~ 'real',
                                                    TRUE ~ 'boot')) %>% 
  group_by(Type, motif_alt_id, Boot) %>% 
  summarise(Mean=mean(Count)) %>% 
  spread(Boot, Mean) %>%
  ungroup() %>% 
  mutate(FC = real / boot, Type = factor(Type, levels=c('iPSC','RFP','GFP'))) %>% 
  arrange(-FC) %>% 
  left_join(tf_motif) %>% 
  select(-boot, -real) 

delta_motif_FC %>% ggplot(aes(y=Type, x=FC)) + geom_density_ridges() + theme_ridges()
```
# Distribution of *relative* TFBS/motif enrichment *between* GFP <-> RFP
```{r}
FC_motif_FC <- all %>% mutate(Type = case_when(grepl('GFP', sample) ~ 'GFP',
                                                    grepl('RFP', sample) ~ 'RFP',
                                                    TRUE ~ 'iPSC'),
                                   Boot = case_when(bootstrap == 'real' ~ 'real',
                                                    TRUE ~ 'boot')) %>% 
  # collapse by Type, motif, and whether bootstrap or real
  group_by(Type, motif_alt_id, Boot) %>% 
  summarise(Mean=mean(Count)) %>% 
  # go wide
  spread(Boot, Mean) %>%
  # so you can calculate FC
  mutate(FC = real / boot) %>% 
  arrange(-FC) %>% 
  # add motif name
  left_join(tf_motif) %>% 
  select(-boot, -real) %>%
  # spread again to allow for GFP vs RFP or iPSC comparisons
  spread(Type, FC) %>% 
  filter(GFP > 1.2) %>% 
  mutate(`FC^2 GFP <-> iPSC` = GFP / iPSC, `FC^2 GFP <-> RFP` = GFP / RFP) %>% 
  arrange(-`FC^2 GFP <-> RFP`) 

plot1 <- FC_motif_FC %>% 
  ggplot(aes(x=`FC^2 GFP <-> RFP`)) + 
  geom_density(fill='gray') + 
  theme_minimal() + 
  geom_vline(xintercept = 1.2, color='blue')
plot1
```

# Distribution of *relative* TFBS/motif enrichment *between* GFP <-> iPSC
```{r}
plot2 <- FC_motif_FC %>% 
  ggplot(aes(x=`FC^2 GFP <-> iPSC`)) + 
  geom_density(fill='gray') + 
  theme_minimal()+ 
  geom_vline(xintercept = 1.2, color = 'blue')
plot2 
```

## Both together
```{r}
plot_grid(plot1 + coord_cartesian(xlim=c(0.5,5.5)), plot2 + coord_cartesian(xlim=c(0.5,5.5)), nrow = 2, align = 'hv', axis='tblr')
```


# What TFBS/motifs are most enriched between...?


## Motifs/TFBS higher in GFP than RFP 
```{r}
#all<-left_join(sample_motifs %>% group_by(sample, motif_alt_id) %>% summarise(Count=n()) %>% ungroup(), bootstrap_counts %>% group_by(motif_alt_id, sample) %>% summarise(Mean = mean(Count), TF = paste(unique(TF), collapse=',')))

 
FC_motif_FC %>% filter(GFP > 1.2) %>% head(1000) %>% DT::datatable()                                          
```

## Motifs/TFBS higher in RFP than GFP 
```{r}
FC_motif_FC %>% filter(GFP > 1.2) %>% arrange(`FC^2 GFP <-> RFP`) %>% head(100) %>% DT::datatable()                                          
```

## Plots of the top motifs enriched in GFP relative to RFP
```{r, fig.height=3, fig.width=6}
plots <- list()
for (i in FC_motif_FC %>% filter(GFP > 1.2) %>% head(n=6) %>% pull(motif_alt_id)){
  plots[[i]] <- plotter(i, scale=T)
}


plot_grid(plotlist = plots)
```

# Gene <-> motif matching
```{r}
# load gtf to map transcript to gene_name

gtf <- read_tsv("/Volumes/data/genomes/1000G_phase2_GRCh37/gencode.v28lift37.metadata.HGNC.gz", col_names = c('Transcript','Gene'))
# gtf <- gtf %>% rowwise() %>% mutate(transcript = ((str_split(V9, ';')[[1]] %>% str_extract('transcript_id.*') %>% na.omit())[1] %>% str_split(., '\"'))[[1]][2],
#                              gene = ((str_split(V9, ';')[[1]] %>% str_extract('gene_name.*') %>% na.omit())[1] %>% str_split(., '\"'))[[1]][2])
```
## Sample data
### Closest TSS
```{r}
samples_full = list.files('/Volumes/data/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/closest_TSS_motifs/', 
                          pattern = '*dat.gz', full.names = T)

closestTSS_data <- list()
for (i in samples_full){
  sample <- str_split(str_split(i, '/')[[1]][10], '\\.')[[1]][1]
  input <- read_tsv(i, col_names = c('sequence_name', 'start', 'end', 'motif', 'fimo_pvalue', 'strand', 'tss_seq','tss_start','tss_end', 'transcript',  'blank','tss_strand','coord1','coord2','blank2','exon_num','size','exon_pos','distance'))
  input <- input %>% mutate(sequence_name = as.character(sequence_name), 
                            tss_seq = as.character(tss_seq), 
                            coord1=as.numeric(coord1), 
                            coord2=as.numeric(coord2), 
                            blank2 = as.character(blank2), 
                            exon_num=as.integer(exon_num),
                            size=as.numeric(size)) %>% mutate(sample = sample)
  closestTSS_data[[sample]] <- input
}
sample_closestTSS = bind_rows(closestTSS_data) %>% mutate(sample=as.factor(sample)) %>% rowwise() %>% mutate(Transcript = str_split(transcript, '_')[[1]][1])

both <- left_join(sample_closestTSS, gtf) %>% 
  filter(!is.na(Gene)) %>% 
  mutate(motif_loc = paste(sequence_name, start, end, sep='_'))

both %>% 
  # remove any TSS over 500kb away
  filter(distance < 500000) %>% 
  # only keep one gene per motif
  group_by(motif_loc, sample, Gene) %>% 
  top_n(1, distance) %>% 
  ungroup() %>% 
  # keep up to two genes per motif
  group_by(motif_loc, sample) %>% 
  top_n(2, distance) %>% 
  ungroup() %>% 
  # arrange by genes most linked to motif  
  group_by(Gene, sample) %>% 
  summarise(Count=n(), paste(motif_loc, collapse=', ')) %>% 
  arrange(-Count)
```

What genes have more PAX6 'associated' motifs compared from GFP to RFP
```{r}
enriched_genes <- both %>% 
  # only keep one gene per motif
  group_by(motif_loc, sample, Gene) %>% 
  top_n(1, distance) %>% 
  ungroup() %>% 
  # keep up to two genes per motif
  group_by(motif_loc, sample) %>% 
  top_n(2, distance) %>% 
  ungroup() %>% 
  # arrange by genes most linked to motif  
  group_by(Gene, sample) %>% 
  summarise(Count=n(), paste(motif_loc, collapse=', ')) %>% 
  ungroup() %>% 
  # collapse to GFP/RFP/iPSC
  mutate(type = case_when(grepl('GFP', sample) ~ 'GFP',
                          grepl('RFP', sample) ~ 'RFP',
                          TRUE ~ 'iPSC')) %>% 
  group_by(Gene, type) %>% 
  summarise(Total=sum(Count)) %>% 
  arrange(-Total) %>% 
  spread(Gene, Total) %>% t() 

colnames(enriched_genes) <- enriched_genes[1,]
enriched_genes <- enriched_genes[-1,] %>% data.frame() %>% rownames_to_column('Gene') %>%  mutate(GFP = as.numeric(GFP), iPSC = as.numeric(iPSC), RFP = as.numeric((RFP)))

enriched_genes[is.na(enriched_genes)] <- 0

enriched_genes %>% mutate(`deltaGFP <-> RFP` = GFP - RFP) %>% arrange(-`deltaGFP <-> RFP`, GFP) %>% head(1000) %>% DT::datatable(rownames = F)

```

## does PAX6 regulate PAX6?
Yes, yes it does.

GFP specific!
```{r}
both %>% 
  # only keep one gene per motif
  group_by(motif_loc, sample, Gene) %>% 
  top_n(1, distance) %>% 
  ungroup() %>% 
  # keep up to two genes per motif
  group_by(motif_loc, sample) %>% 
  top_n(2, distance) %>% 
  ungroup() %>% 
  # arrange by genes most linked to motif  
  group_by(Gene, sample) %>% 
  summarise(Count=n(), `Motif Locations` = paste(motif_loc, collapse=', ')) %>% 
  arrange(-Count) %>% 
  filter(Gene=='PAX6')
```

```
