---
title: Motif / TFBS Stats
author: David McGaughey
date: '`r format(Sys.Date(), "%Y-%m-%d")`'
output: 
  html_notebook:
    theme: flatly
    toc: true
    code_folding: hide
---


```{r, message=F, warning=F, results='hide'}
# Load Libraries without printing any warnings or messages
library(tidyverse)
library(ggridges)
library(cowplot)
library(DESeq2)
library(ggrepel)
```
```{r}
# load precomputed data from src/motif_heavy_lift.R
load('/Volumes/data/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/Rdata/resLFC.Rdata')
load('/Volumes/data/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/Rdata/dds.Rdata')
load('/Volumes/data/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/Rdata/sample_bootstrap_counts.Rdata')
load('/Volumes/data/projects/nei/hufnagel/iPSC_RPE_ATAC_Seq/Rdata/sample_closestTSS.Rdata')
load('~/git/ipsc_rpe_atac/data/tf_motif.Rdata')
```

# Motif/TF assessed
There are ~3,000 in cis-bp

I've filtered it down to ~1,000 by only including TF that are log2FC > 1 between RPE <-> iPSC or GFP+ <-> RFP+

```{r, results = 'hide'}
tf_motif %>% DT::datatable()
```

```{r}
# motif ggridges plotter
# removes GFP 
plotterGi <- function(motif, scale=TRUE) {
  if (scale == TRUE){
    sample_bootstrap_counts <- sample_bootstrap_counts %>% filter(motif_alt_id == !!motif) %>% 
      filter(!grepl('RFP',sample)) %>% 
      group_by(sample, motif_alt_id) %>% 
      mutate(`Z score` = scale(Count)) %>% 
      ungroup() 
    sample_bootstrap_counts %>% 
      filter(!grepl('RFP',sample)) %>% 
      #mutate(sample = factor(sample, levels=c('iPSC_IIi_9','iPSC_IIJ_10','iPSC_IIK_11','iPSC_IIL_12','RFP_IIE_1','RFP_IIF_2','RFP_IIG_3','RFP_IIH_4','GFP_IIE_5','GFP_IIF_6','GFP_IIG_7','GFP_IIH_8'))) %>% 
      mutate(sample = factor(sample, levels=c('iPSC_IIi_9','iPSC_IIJ_10','iPSC_IIK_11','iPSC_IIL_12','GFP_IIE_5','GFP_IIF_6','GFP_IIG_7','GFP_IIH_8'))) %>% 
      ggplot(aes(x=`Z score`, y=sample)) + 
      geom_density_ridges() +
      geom_point(data = sample_bootstrap_counts %>% 
                   filter(bootstrap == 'real', motif_alt_id == !!motif) %>% 
                   ungroup(), aes(x=`Z score`,y=sample), colour='blue', size=2, alpha=0.5) +
      scale_y_discrete(expand = c(0.01, 0)) +
      theme_ridges() + 
      ggtitle(paste0(motif, ' (', 
                     tf_motif %>% filter(motif_alt_id == !!motif) %>% pull(TF),
                     ')'))
  }
  else {
    sample_bootstrap_counts %>% filter(motif_alt_id == !!motif) %>% 
      filter(!grepl('RFP',sample)) %>% 
      #mutate(sample = factor(sample, levels=c('iPSC_IIi_9','iPSC_IIJ_10','iPSC_IIK_11','iPSC_IIL_12','RFP_IIE_1','RFP_IIF_2','RFP_IIG_3','RFP_IIH_4','GFP_IIE_5','GFP_IIF_6','GFP_IIG_7','GFP_IIH_8'))) %>% 
      mutate(sample = factor(sample, levels=c('iPSC_IIi_9','iPSC_IIJ_10','iPSC_IIK_11','iPSC_IIL_12','GFP_IIE_5','GFP_IIF_6','GFP_IIG_7','GFP_IIH_8'))) %>% 
      ggplot(aes(x=Count, y=sample)) + 
      geom_density_ridges() +
      geom_point(data = sample_motifs %>% 
                   filter(motif_alt_id == !!motif) %>% 
                   group_by(sample, motif_alt_id) %>%
                   summarise(Count=n()) %>% ungroup(), aes(x=Count,y=sample), colour='blue', size=2, alpha=0.5) +
      scale_y_discrete(expand = c(0.01, 0)) +
      theme_ridges() + 
      ggtitle(paste0(motif, ' (', 
                     tf_motif %>% filter(motif_alt_id == !!motif) %>% pull(TF),
                     ')'))
  }
}
```
# Plot enrichment  specific motif/TFBS

## TET1
```{r}
plotterGi('M0610_1.02')
```

## DNMT1
```{r}
plotterGi('M0609_1.02')
```

## PAX6
```{r}
plotterGi('M6410_1.02')
```

## SOX10
```{r}
plotterGi('M6470_1.02')
```

## MITF
```{r}
plotterGi('M6345_1.02')
```

## HES1
```{r}
plotterGi('M6271_1.02')
```

## SIX3
```{r}
plotterGi('M1016_1.02')
```

## SIX2
```{r}
plotterGi('M0952_1.02')
```

# Distribution of TFBS/motif enrichment
## Fold Change
```{r}
delta_motif_FC <- sample_bootstrap_counts %>% mutate(Type = case_when(grepl('GFP', sample) ~ 'GFP',
                                                    grepl('RFP', sample) ~ 'RFP',
                                                    TRUE ~ 'iPSC'),
                                   Boot = case_when(bootstrap == 'real' ~ 'real',
                                                    TRUE ~ 'boot')) %>% 
  group_by(Type, motif_alt_id, Boot) %>% 
  mutate(`Z score` = scale(Count)) %>% 
  summarise(Mean=mean(Count)) %>% 
  spread(Boot, Mean) %>%
  ungroup() %>% 
  mutate(FC = real / boot, Type = factor(Type, levels=c('iPSC','RFP','GFP'))) %>% 
  arrange(-FC) %>% 
  left_join(tf_motif) %>% 
  select(-boot, -real) 

delta_motif_FC %>% ggplot(aes(y=Type, x=log2(FC))) + geom_density_ridges() + theme_ridges() 

delta_motif_FC %>% ggplot(aes(y=Type, x=FC)) + geom_density_ridges() + theme_ridges() 
```



```{r}
# calculate z scores for sample_bootstrap_counts
Z_scoring <- sample_bootstrap_counts %>% 
  # collapse by Type, motif, and whether bootstrap or real
  group_by(sample, motif_alt_id) %>% 
  mutate(`Z score` = scale(Count)) %>% 
  filter(bootstrap == 'real') %>% 
  select(-bootstrap, -TF) %>% 
  left_join(tf_motif)

#Z_scoring %>% head()
```
# Distribution of *relative* TFBS/motif enrichment *between* GFP <-> RFP
```{r}
FC_motif_FC <- sample_bootstrap_counts %>% mutate(Type = case_when(grepl('GFP', sample) ~ 'GFP',
                                                    grepl('RFP', sample) ~ 'RFP',
                                                    TRUE ~ 'iPSC'),
                                   Boot = case_when(bootstrap == 'real' ~ 'real',
                                                    TRUE ~ 'boot')) %>% 
  # collapse by Type, motif, and whether bootstrap or real
  group_by(Type, motif_alt_id, Boot) %>% 
  summarise(Mean=mean(Count)) %>% 
  # go wide
  spread(Boot, Mean) %>%
  # so you can calculate FC
  mutate(FC = real / boot) %>% 
  arrange(-FC) %>% 
  # add motif name
  left_join(tf_motif) %>% 
  select(-boot, -real) %>%
  # spread again to sample_bootstrap_countsow for GFP vs RFP or iPSC comparisons
  spread(Type, FC) %>% 
  filter(GFP > 1.2) %>% 
  mutate(`FC^2 GFP <-> iPSC` = GFP / iPSC, `FC^2 GFP <-> RFP` = GFP / RFP) %>% 
  arrange(-`FC^2 GFP <-> RFP`) 

plot1 <- FC_motif_FC %>% 
  ggplot(aes(x=`FC^2 GFP <-> RFP`)) + 
  geom_density(fill='gray') + 
  theme_minimal() + 
  geom_vline(xintercept = 1.2, color='blue')
plot1
```

# Distribution of *relative* TFBS/motif enrichment *between* GFP <-> iPSC
```{r}
plot2 <- FC_motif_FC %>% 
  ggplot(aes(x=`FC^2 GFP <-> iPSC`)) + 
  geom_density(fill='gray') + 
  theme_minimal()+ 
  geom_vline(xintercept = 1.2, color = 'blue')
plot2 
```

## Both together
```{r}
plot_grid(plot1 + coord_cartesian(xlim=c(0.5,5.5)), plot2 + coord_cartesian(xlim=c(0.5,5.5)), nrow = 2, align = 'hv', axis='tblr')
```
